A GRASP algorithm for the concrete delivery problem

### 2024-03-25 ArticleCDP/original/articleCDP.tex
correct indices in model
last version

\documentclass{article}
% \documentclass[review]{elsarticle}
\usepackage{epsfig}
\usepackage{amsmath}
\usepackage{mathtools}
\usepackage{algorithm}
\usepackage[algo2e,noend]{algorithm2e}
\usepackage[compatible,noend]{algpseudocode}
\usepackage{enumitem}
\usepackage{xspace}
\usepackage[table,xcdraw]{xcolor}
\usepackage{amsfonts}
\usepackage{pifont}
\usepackage{bbding}
\usepackage[english]{babel}
\usepackage{adjustbox}
\usepackage{tikz}
\usepackage{subcaption}
\usetikzlibrary{automata, topaths, calc, positioning, shapes, backgrounds, fit, matrix}
\usepackage{caption}
\usepackage{tabularx}
\usepackage[T1]{fontenc}
\usepackage{graphicx}
\usepackage{natbib}
\usepackage{datatool}
\usepackage{array}
\usepackage{pgfplotstable}
\pgfplotsset{width=7cm,compat=newest}
\usepackage{booktabs}
\usepackage{longtable}
\usepackage{pdflscape}
\usepackage{afterpage}
\usepackage{capt-of}
\usepackage{multirow}
\usepackage{float}
\usepackage[text={15cm,21cm},centering]{geometry}
\usepackage{setspace}
\usepackage{array}
\usepackage{lineno}
\usepackage{breqn}
\usepackage{tkz-graph}
\usepackage{varwidth}
\usepackage{url}
\allowdisplaybreaks
\newcolumntype{H}{>{\setbox0=\hbox\bgroup}c<{\egroup}@{}}

@@

\begin{table}[!ht]
    \centering
    \caption{Information of the illustration instance}
    \label{tab:instance_detail}
    \resizebox*{0.8\textwidth}{!}{
        \begin{tabular}{@{}c|cccccccccccccc@{}}
            \toprule
            Order         & 1     & 2     & 3     & 4     & 5     & 6     & 7     & 8       \\ \midrule
            Customer      & 1     & 2     & 3     & 4     & 4     & 5     & 5     & 5       \\
            \midrule
            Demand   & 24   & 32   & 36   & 40   & 16  & 20  & 32  & 8 \\
            \midrule
            Plant    & 1     & 2     & 1     & 2     & 2    & 2  &  2 & 2     \\
            \midrule
            TW       & 08:45 & 08:00 & 07:00 & 07:00 & 07:00 & 06:50 & 06:50 & 06:50  \\ \bottomrule
        \end{tabular}
    }
\end{table}

\begin{figure}[htb]
    \centering
    \caption{Gantt chart of the loading and unloading operations of the illustration instance}
    \label{fig:plants_schedules}
        \includegraphics[width=\linewidth]{exemple.pdf}
\end{figure}

@@

    Go to line 6 \tcp{backtrack to $j'$ position and retry to insert $D^{j'}_{o}$}
    }
    \uElse{
     Search for a precedent delivery node $D^{j'}_o, j'<j$ with $M[D^{j'}_o]\neq \emptyset$

@@

                }
            }
        }
        \KwRet{$Sol$}
    \end{algorithm}

When confronted with scenarios where several customers require their services to start simultaneously or within the same time window, we observed that with Algorithm~\ref{alg:greedyCustIns}, some customers remain unscheduled when resources are limited. To address this limitation, we introduce the delivery-based insertion heuristic.

@@

Between lines $14$ and $18$, for each delivery $d$ in $CL'$, the algorithm determines the best available driver $k$ who can visit it. Then, the corresponding pair of loading and unloading tasks $(L^k_d, U^k_d)$, along with its costs, are stored in $M[d]$. This is done using the $SimulateVisit$ procedure described earlier. Then, the algorithm constructs at line $20$ a list $T$ containing all pairs of loading and unloading tasks for the delivery nodes in $CL'$.

If $T$ is empty, the algorithm assumes that the customers associated with the nodes in $CL'$ cannot be scheduled, and it removes those customers' existing schedules from the solution. It then excludes $CL'$ from $CL$ at line $25$ and proceeds to the next iteration.

Otherwise, from line $26$ to line $29$, the algorithm randomly selects from $T$ a pair of loading and unloading tasks whose cost falls within a specified range determined by the parameter $\theta$ and inserts this pair into $Sol$. The algorithm then updates the remaining orders and customer quantities and excludes the scheduled delivery node from $CL$. If an order $o$ is not completely fulfilled after visiting $d^j_{o}$, the next delivery node $d^{j+1}_{o}$ is added to $CL$.  Otherwise, if the customer has remaining orders, the first delivery node $d^{1}_{o'}$ of another randomly selected order $o'$ is added to $CL$.

Unlike the customer-based insertion, the delivery-based insertion allows each customer to have one of its delivery nodes considered for scheduling at each iteration.
%In these two heuristics, the greedy aspect of GRASP refers to the creation of the set $RCL$, the probabilistic aspect is the random selection in $RCL$, and the adaptive part is the updating of $CL$ and the re-evaluation of the costs.

@@

    $M \leftarrow \emptyset $ \tcp{List of pairs of loading -- unloading tasks for each delivery node}

@@

                    $best\_cost \leftarrow c$
                }
            }

@@

            $Sol \leftarrow Sol \backslash \{(L^k_j,U^k_j), \forall j \in D_c \}$
           }
           $CL \leftarrow CL\backslash CL'$ 
        }
        \uElse{
            $c_{min} \leftarrow \min\left\{ cost \ | \ (L^k_{d}, U^k_{d},cost,k) \in T   \right\} $

            $c_{max} \leftarrow \max\left\{ cost \ | \ (L^k_{d}, U^k_{d},cost,k) \in T   \right\} $

@@

We also implemented two moves based on the relocate operator. The Consolidate Load operator is applied on an order with a remaining demand between $]Q_{min},Q_{max}]$. If this remaining demand is served with two drivers with capacity $Q_{min}$, then the operator tries to consolidate it to be served by a single driver with a capacity of $Q_{max}$. Next, the Relocate Driver operator removes a delivery node from its current route and inserts it in another driver's route. This enables the redistribution of delivery nodes among drivers, potentially reducing drivers underutilization and overtime costs.

\section{Computational experiments}
\label{sec:cdp_comp_exp}

@@

    \begin{minipage}{12cm}

@@

    {\footnotesize $TC$: travel cost, $UQ$: undelivered quantity, $FDD$: first delivery delay, $mFDD$: maximum of first delivery delays, $DUC$: driver underutilization cost, $DOC$: driver overtime cost.}        

@@

\begin{table}[htb]
    \centering
    \caption{RMC delivery completion within different runtimes of the GRASP}
    \label{tab_cdp_runtime}
    \scriptsize
    \resizebox*{0.7\textwidth}{!}{
    \begin{tabular}{llrrrrr}
    \toprule
     &  & \multicolumn{5}{c}{Runtime (min)} \\
     \cmidrule{3-7}
     &  & 1 & 5 & 10 & 30 & 60 \\
    \midrule
    Small & \%Load & 100.0 & 100.0 & 100.0 & 100.0 & 100.0 \\
Medium & \%Load & 98.9 & 99.3 & 99.4 & 99.5 & 99.5 \\
Large & \%Load & 98.8 & 99.3 & 99.4 & 99.6 & 99.6 \\
    \bottomrule
    \end{tabular}
    }
\end{table}

The medium instance $C\_42\_47\_104\_8$ has an unserved customer with a demand of 12 $m^3$ and a due time of $07\text{:}00$. Upon investigation, we find that even though we have $104$ drivers in this instance, the number of drivers who start their shift before $07\text{:}00$ is insufficient to serve this customer within the 60-minute window we defined. Thus, the value of $UQ$ that we report in Table~\ref{tab:cdp_all_instances} appears optimal given the driver shift schedule given as input to our algorithm. We confirmed this observation by removing the driver start shift and overtime constraints and resolving our instances five more times. All orders are now fully serviced on each run, as shown in Table~\ref{tab:cdp_all_instances_no_shift}. Table~\ref{tab:cdp_lift_shift_constraints} compares the averages of the results from Tables \ref{tab:cdp_all_instances} and \ref{tab:cdp_all_instances_no_shift}. We found that the maximum delay dropped to zero for small instances, and for medium (large) instances, it was reduced by almost 60\% (40\%) on average. This suggests that when we input a good work schedule, our algorithm works effectively to ensure that all orders are filled and that delays for first deliveries and travel costs are minimized. We can see that the average travel time decreases in all cases except the large one, where the difference is not significant.

@@

        {\footnotesize $TC$: travel cost, $UQ$: undelivered quantity, $FDD$: first delivery delay, $mFDD$: maximum of first delivery delays.}        

@@

\begin{table}[!ht]
    \centering
    \small
    \caption{Comparison of the average results with and without shift and overtime constraints}
    \label{tab:cdp_lift_shift_constraints}
    \resizebox{0.8\textwidth}{!}{
    \begin{tabular}{lrrrrrrrr}
    \toprule
     & \multicolumn{4}{c}{Default parameters}&\multicolumn{4}{c}{$\mu=0$, $\eta,\varLambda=\infty$, $h_k=0$} \\
     \cmidrule(l){2-5} \cmidrule(l){6-9}
    &   \multicolumn{1}{c}{$TC$}& \multicolumn{1}{c}{$UQ$} & \multicolumn{1}{c}{$FDD$} & \multicolumn{1}{c}{$mFDD$}  & \multicolumn{1}{c}{($TC$)} & \multicolumn{1}{c}{$UQ$} & \multicolumn{1}{c}{$FDD$} & \multicolumn{1}{c}{$mFDD$}  \\
    & \multicolumn{1}{c}{(min)} & \multicolumn{1}{c}{(m$^3$)} & \multicolumn{1}{c}{(min)} & \multicolumn{1}{c}{(min)} & \multicolumn{1}{c}{(min)} & \multicolumn{1}{c}{(m$^3$)} & (min) & \multicolumn{1}{c}{(min)} \\
    \midrule
    Small & 1,933.6 & 0.0 & 7.4 & 4.5 & 1,822.7 & 0.0 & 0.0 & 0.0 \\
Medium & 15,070.0 & 9.9 & 403.0 & 46.9 & 14,719.3 & 0.0 & 68.5 & 18.4 \\
Large & 21,810.5 & 14.7 & 945.0 & 56.4 & 22,283.6 & 0.0 & 168.7 & 31.0 \\
    \bottomrule
    \end{tabular}
    }
    \begin{minipage}{12cm}
        \vspace{0.1cm}
        {\footnotesize $TC$: travel cost, $UQ$: undelivered quantity, $FDD$: first delivery delay, $mFDD$: maximum of first delivery delays.}        
        \end{minipage}
\end{table}

@@

\begin{table}[!ht]
    \centering
    \caption{Trade-off between travel cost and first delivery delay}
    \label{tab:cdp_tradeoff}
    \begin{adjustbox}{max width=0.8\textwidth}
    \begin{tabular}{lcccccc}
    \toprule
     & \multicolumn{3}{c}{Minimize TC} & \multicolumn{3}{c}{Minimize $FDD$} \\
     \cmidrule(l){2-4} \cmidrule(l){5-7}
     & $TC$ & $FDD$ & $mFDD$ & $TC$ & $FDD$ & $mFDD$ \\
     & (min) & (min) & (min) & (min) & (min) & (min) \\
    \midrule
    Small & 1,821.3 & 0.0 & 0.0 & 1,831.6 & 0.0 & 0.0 \\
    Medium & 14,311.5 & 89.6 & 23.0 & 14,679.3 & 46.9 & 13.4 \\
    Large & 21,403.8 & 170.5 & 31.4 & 22,194.7 & 101.7 & 29.4 \\
    \bottomrule
    \end{tabular}
\end{adjustbox}

@@

\vspace{1.5cm} \noindent \textbf{Acknowledgments}

@@

\bibliographystyle{plainnat}
\bibliography{references}

>>>>

\documentclass[]{article}
\usepackage{epsfig}
\usepackage{amsmath}
\usepackage{mathtools}
\usepackage{algorithm}
\usepackage[algo2e,noend]{algorithm2e}
\usepackage[compatible,noend]{algpseudocode}
\usepackage{enumitem}
\usepackage{xspace}
\usepackage[table,xcdraw]{xcolor}
\usepackage{amsfonts}
\usepackage{pifont}
\usepackage{bbding}
\usepackage[english]{babel}
\usepackage{adjustbox}
\usepackage{tikz}
\usepackage{subcaption}
\usetikzlibrary{automata, topaths, calc, positioning, shapes, backgrounds, fit, matrix}
\usepackage{caption}
\usepackage{tabularx}
\usepackage[T1]{fontenc}
\usepackage{graphicx}
\usepackage{natbib}
\usepackage{datatool}
\usepackage{array}
\usepackage{pgfplotstable}
\pgfplotsset{width=7cm,compat=newest}
\usepackage{booktabs}
\usepackage{longtable}
\usepackage{pdflscape}
\usepackage{afterpage}
\usepackage{capt-of}
\usepackage{multirow}
\usepackage{float}
\usepackage[text={15cm,21cm},centering]{geometry}
\usepackage{setspace}
\usepackage{array}
\usepackage{lineno}
\usepackage{breqn}
\usepackage{tkz-graph}
\usepackage{varwidth}
\usepackage{url}
\allowdisplaybreaks
\newcolumntype{H}{>{\setbox0=\hbox\bgroup}c<{\egroup}@{}}

@@

{
    % \small
\begin{table}[!ht]
    \centering
    \caption{Information of the illustration instance}
    \label{tab:instance_detail}
    \resizebox*{0.8\textwidth}{!}{
        \begin{tabular}{@{}c|cccccccccccccc@{}}
            \toprule
            Order         & 1     & 2     & 3     & 4     & 5     & 6     & 7     & 8       \\ \midrule
            Customer      & 1     & 2     & 3     & 4     & 4     & 5     & 5     & 5       \\
            \midrule
            Demand   & 24   & 32   & 36   & 40   & 16  & 20  & 32  & 8 \\
            \midrule
            Plant    & 1     & 2     & 1     & 2     & 2    & 2  &  2 & 2     \\
            \midrule
            TW       & 08:45 & 08:00 & 07:00 & 07:00 & 07:00 & 06:50 & 06:50 & 06:50  \\ \bottomrule
        \end{tabular}
    }
\end{table}
}
\begin{figure}[htb]
    \centering
    \caption{Gantt chart of the loading and unloading operations of the illustration instance}
    \label{fig:plants_schedules}
        \includegraphics[width=\linewidth]{gantt_example.pdf}
\end{figure}

@@

    Go to line 6 \small \tcp{backtrack and retry to insert $D^{j'}_{o}$}
    }
    \uElse{
     Search for a precedent delivery node $D^{j'}_o, j'<j$ with $M[D^{j'}_o]\neq \emptyset$

@@

                }
            }
        }
        \KwRet{$Sol$}
    \end{algorithm}
}
When confronted with scenarios where several customers require their services to start simultaneously or within the same time window, we observed that with Algorithm~\ref{alg:greedyCustIns}, some customers remain unscheduled when resources are limited. To address this limitation, we introduce the delivery-based insertion heuristic, which allows each customer to have one of its delivery nodes considered for scheduling at each iteration.

@@

Between lines $14$ and $18$, for each delivery $d$ in $CL'$, the algorithm determines the best available driver $k$ who can visit it. Then, the corresponding pair of loading and unloading tasks $(L^k_d, U^k_d)$, along with its costs, are stored in $M[d]$. This is done using the $SimulateVisit$ procedure described earlier. Then, the algorithm constructs at line $20$ a list $T$ containing all pairs of loading and unloading tasks for the delivery nodes in $CL'$.
If $T$ is empty, the algorithm assumes that the customers associated with the nodes in $CL'$ cannot be scheduled, and it removes those customers' existing schedules from the solution. It then excludes $CL'$ from $CL$ at line $25$ and proceeds to the next iteration.

Otherwise, from line $26$ to line $29$, the algorithm randomly selects from $T$ a pair of loading and unloading tasks whose cost falls within a specified range determined by the parameter $\theta$ and inserts this pair into $Sol$. The algorithm then updates the remaining orders and customer quantities and excludes the scheduled delivery node from $CL$. If an order $o$ is not completely fulfilled after visiting $d^j_{o}$, the next delivery node $d^{j+1}_{o}$ is added to $CL$.  Otherwise, if the customer has remaining orders, the first delivery node $d^{1}_{o'}$ of another randomly selected order $o'$ is added to $CL$.
%In these two heuristics, the greedy aspect of GRASP refers to the creation of the set $RCL$, the probabilistic aspect is the random selection in $RCL$, and the adaptive part is the updating of $CL$ and the re-evaluation of the costs.

@@

    $M \hspace{-2mm} \leftarrow \emptyset $ \small \tcp{List of pairs of loading -- unloading tasks for each delivery node}

@@

                    $best\_cost \leftarrow cost$
                }
            }

@@

            $Sol \leftarrow Sol \backslash \{(L^k_j,U^k_j), \forall j \in D_c \}$
           }
           $CL \leftarrow CL\backslash CL'$ 
        }
        \uElse{
            $c_{min} \leftarrow \min\left\{ cost \ | \ (L^k_{d}, U^k_{d},cost,k) \in T   \right\} $, $c_{max} \leftarrow \max\left\{ cost \ | \ (L^k_{d}, U^k_{d},cost,k) \in T   \right\} $

@@

We also implemented two moves based on the relocate operator. The Consolidate Load operator is applied on an order with a remaining demand between $]Q_{min},Q_{max}]$. If this remaining demand is served with two drivers with capacity $Q_{min}$, then the operator tries to consolidate it to be served by a single driver with a capacity of $Q_{max}$. Next, the Relocate Driver operator removes a delivery node from its current route and inserts it in another driver's route. This enables the redistribution of delivery nodes among drivers, potentially reducing drivers underutilization and overtime costs.
% \clearpage
\section{Computational experiments}
\label{sec:cdp_comp_exp}

@@

    \begin{minipage}{\textwidth}

@@

    {\scriptsize \hspace*{-5mm} $TC$: travel cost, $UQ$: undelivered quantity, $FDD$: first delivery delay, $mFDD$: maximum of first delivery delays, $DUC$: driver underutilization cost, $DOC$: driver overtime cost.}        

@@

\begin{table}[ht!]
    \centering
    \caption{RMC delivery completion within different runtimes of the GRASP}
    \label{tab_cdp_runtime}
    \scriptsize
    \resizebox*{0.7\textwidth}{!}{
    \begin{tabular}{llrrrrr}
    \toprule
    %  &  & \multicolumn{5}{c}{Runtime (min)} \\
    %  \cmidrule{3-7}
     &  & \multicolumn{1}{c}{1 min} & \multicolumn{1}{c}{5 min} & \multicolumn{1}{c}{10 min} & \multicolumn{1}{c}{30 min} &\multicolumn{1}{c}{60 min} \\
    \midrule
    Small & \%Load & 100.0 & 100.0 & 100.0 & 100.0 & 100.0 \\
Medium & \%Load & 98.9 & 99.3 & 99.4 & 99.5 & 99.5 \\
Large & \%Load & 98.8 & 99.3 & 99.4 & 99.6 & 99.6 \\
    \bottomrule
    \end{tabular}
    }
\end{table}

The medium instance $C\_42\_47\_104\_8$ has an unserved customer with a demand of 12 $m^3$ and a due time of $07\text{:}00$. Upon investigation, we find that even though we have $104$ drivers in this instance, the number of drivers who start their shift before $07\text{:}00$ is insufficient to serve this customer within the 60-minute window we defined. Thus, the value of $UQ$ that we report in Table~\ref{tab:cdp_all_instances} appears optimal given the driver shift schedule provided as input to our algorithm. We confirmed this observation by removing the driver start shift and overtime constraints and resolving our instances five more times. All orders are now fully serviced on each run, as shown in Table~\ref{tab:cdp_all_instances_no_shift}. Table~\ref{tab:cdp_lift_shift_constraints} compares the averages of the results from Tables \ref{tab:cdp_all_instances} and \ref{tab:cdp_all_instances_no_shift}. We found that the maximum delay dropped to zero for small instances, and for medium (large) instances, it was reduced by almost 60\% (40\%) on average. This suggests that when we input a good work schedule, our algorithm works effectively to ensure that all orders are filled and that delays for first deliveries and travel costs are minimized. We can see that the average travel time decreases in all cases except the large one, where the difference is not significant.

@@

        {\scriptsize $TC$: travel cost, $UQ$: undelivered quantity, $FDD$: first delivery delay, $mFDD$: maximum of first delivery delays.}        

@@

\begin{table}[!ht]
    \centering
    \small
    \caption{Comparison of the average results with and without shift and overtime constraints}
    \label{tab:cdp_lift_shift_constraints}
    \resizebox{0.8\textwidth}{!}{
    \begin{tabular}{lrrrrrrrr}
    \toprule
     & \multicolumn{4}{c}{Default parameters}&\multicolumn{4}{c}{$\mu=0$, $\eta,\varLambda=\infty$, $h_k=0$} \\
     \cmidrule(l){2-5} \cmidrule(l){6-9}
    &   \multicolumn{1}{c}{$TC$}& \multicolumn{1}{c}{$UQ$} & \multicolumn{1}{c}{$FDD$} & \multicolumn{1}{c}{$mFDD$}  & \multicolumn{1}{c}{($TC$)} & \multicolumn{1}{c}{$UQ$} & \multicolumn{1}{c}{$FDD$} & \multicolumn{1}{c}{$mFDD$}  \\
    & \multicolumn{1}{c}{(min)} & \multicolumn{1}{c}{(m$^3$)} & \multicolumn{1}{c}{(min)} & \multicolumn{1}{c}{(min)} & \multicolumn{1}{c}{(min)} & \multicolumn{1}{c}{(m$^3$)} & (min) & \multicolumn{1}{c}{(min)} \\
    \midrule
    Small & 1,933.6 & 0.0 & 7.4 & 4.5 & 1,822.7 & 0.0 & 0.0 & 0.0 \\
Medium & 15,070.0 & 9.9 & 403.0 & 46.9 & 14,719.3 & 0.0 & 68.5 & 18.4 \\
Large & 21,810.5 & 14.7 & 945.0 & 56.4 & 22,283.6 & 0.0 & 168.7 & 31.0 \\
    \bottomrule
    \end{tabular}
    }
    \begin{minipage}{\textwidth}
        \vspace{0.1cm}
        {\scriptsize $TC$: travel cost, $UQ$: undelivered quantity, $FDD$: first delivery delay, $mFDD$: maximum of first delivery delays.}        
        \end{minipage}
\end{table}

@@

\begin{table}[!ht]
    \centering
    \caption{Trade-off between travel cost and first delivery delay}
    \label{tab:cdp_tradeoff}
    \begin{adjustbox}{max width=0.9\textwidth}
    \begin{tabular}{lcccccc}
    \toprule
     & \multicolumn{3}{c}{Minimize TC} & \multicolumn{3}{c}{Minimize $FDD$} \\
     \cmidrule(l){2-4} \cmidrule(l){5-7}
     & \multicolumn{1}{c}{$TC$ (min)} & \multicolumn{1}{c}{$FDD$ (min)} & \multicolumn{1}{c}{$mFDD$ (min)} & \multicolumn{1}{c}{$TC$ (min)} & \multicolumn{1}{c}{$FDD$ (min)} & \multicolumn{1}{c}{$mFDD$ (min)}\\
    %  & \multicolumn{1}{c}{(min)} & \multicolumn{1}{c}{(min)}  & \multicolumn{1}{c}{(min)} & \multicolumn{1}{c}{(min)}  & \multicolumn{1}{c}{(min)} & \multicolumn{1}{c}{(min)}  \\
    \midrule
    Small & 1,821.3 & 0.0 & 0.0 & 1,831.6 & 0.0 & 0.0 \\
    Medium & 14,311.5 & 89.6 & 23.0 & 14,679.3 & 46.9 & 13.4 \\
    Large & 21,403.8 & 170.5 & 31.4 & 22,194.7 & 101.7 & 29.4 \\
    \bottomrule
    \end{tabular}
\end{adjustbox}

@@

\vspace{1.cm} \noindent \textbf{Acknowledgments}

@@

 \bibliographystyle{plainnat} 
 \bibliography{References}
